---
title: "Metabolites Figure"
author: "Sam McMillan"
date: "2023-11-29"
output: html_document
---

```{r setup, require=T, include=T}
library(tidyverse)
library(janitor)
library(randomForest)
library(pheatmap)
library(psych)

sessionInfo()
```

# Get Dataframes
Load in the dataframes we made from the "Data Organization and Stats" markdown
```{r}
samplekey<-readRDS("Dataframes/samplekey.rds")
bch_raw<-readRDS("Dataframes/bch_raw.rds")
bch_metadata<-readRDS("Dataframes/bch_metadata.rds")
bch_siglist<-readRDS("Dataframes/Stats/bch_siglist.rds")
tax_raw<-readRDS("Dataframes/tax_raw.rds")
fam_for_cor<-readRDS("Dataframes/fam_for_cor.rds")
bch_for_cor<-readRDS("Dataframes/bch_for_cor.rds")
tax_metadata<-readRDS("Dataframes/tax_metadata.rds")

source("Scripts/Color codes.R")
```

# Random Forest
We'll use random forest to identify important metabolites
```{r}
bch_long<-bch_raw%>%
  gather(key="CLIENT.IDENTIFIER",value="ScaledImpData",-BIOCHEMICAL)%>%
  mutate(CLIENT.IDENTIFIER=factor(CLIENT.IDENTIFIER,levels=levels(samplekey$CLIENT.IDENTIFIER)))%>%
  left_join(.,samplekey,by="CLIENT.IDENTIFIER")

#Random Forest
set.seed(102309)
bch_forest<-bch_raw%>%
  column_to_rownames("BIOCHEMICAL")%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("CLIENT.IDENTIFIER")%>%
  left_join(samplekey,., by="CLIENT.IDENTIFIER")%>%
  select(-CLIENT.IDENTIFIER,-TIME.POINT.2,-SUBJ.ID)%>%
  clean_names()%>%
  randomForest(prepost ~ .,data=., proximity=T, importance=TRUE, ntree=1000)

#Prepare to Plot
bch_forest_plot<-bch_forest$importance%>%
  as.data.frame()%>%
  rownames_to_column("biochem")%>%
  left_join(.,bch_metadata%>%
              select(BIOCHEMICAL,SUPER.PATHWAY,SUB.PATHWAY, dir)%>%
              mutate(biochem=make_clean_names(BIOCHEMICAL)),by="biochem")%>%
  filter(MeanDecreaseAccuracy>0)%>%
  arrange(-MeanDecreaseAccuracy)%>%
  group_by(dir)%>%
  mutate(count=1:n())%>%
  arrange(dir,MeanDecreaseAccuracy)%>%
  mutate(plotnames=str_remove_all(BIOCHEMICAL,"\\*"))%>%
  mutate(plotnames=fct_inorder(plotnames))

#plot
ggplot(data=bch_forest_plot%>%filter(count<=25), aes(y=MeanDecreaseAccuracy,x=plotnames, fill=SUPER.PATHWAY))+
  geom_point(shape=21, size=2)+
  scale_fill_manual(values=anno_color$SUPER.PATHWAY)+
  coord_flip()+
  theme_bw()+
  labs(fill="", y="Mean Decrease Accuracy",x="")+
  theme(axis.text.y=element_text(size=6),
        axis.text.x=element_text(size=6),
        axis.title.y=element_text(size=8),
        legend.position="top",
        legend.text = element_text(size=6),
        legend.key.size = unit(c(0.1,0.2),"in"),
        legend.box.margin=unit(c(0,1.5,-.15,0),"in"),
        plot.margin = unit(c(0,0.25,0,0),"in"))+
  guides(fill=guide_legend(nrow=3))

ggsave(paste0("Figures/Metabolites/metabolite_50_forest.pdf"), height=5, width=3.5, units="in")

##Error Rate
bch_forest

##Check for non significant
print(paste0((bch_forest_plot%>%filter(count<=25))$BIOCHEMICAL[!(bch_forest_plot%>%filter(count<=25))$BIOCHEMICAL %in% bch_siglist$BIOCHEMICAL]," is important by random forest but not significant by linear modelling"))

#Outputs for writing
bch_forest_plot%>%
  group_by(dir)%>%
  summarize(count=n())

bch_forest_plot%>%
  filter(count<=25)%>%
  group_by(SUPER.PATHWAY,dir)%>%
  summarise(count=n())%>%
  spread(key=dir,value=count)%>%
  ungroup()%>%
  mutate(total=rowSums(.[2:3],na.rm=T))%>%
  arrange(-total)

bch_forest_plot%>%
  filter(SUPER.PATHWAY=="Lipid" & count<=25)%>%
  group_by(SUB.PATHWAY,dir)%>%
  summarize(count=n())%>%
  spread(key=dir,value=count)%>%
  ungroup()%>%
  mutate(total=rowSums(.[2:3],na.rm=T))%>%
  arrange(-total)


#make list for filtering
top_bch_forest<-as.vector((bch_forest_plot%>%filter(count<=25))$BIOCHEMICAL)

#Save list for later filtering
saveRDS(top_bch_forest, "Dataframes/top_bch_forest.rds")

#Save for export
saveRDS(bch_forest$importance%>%as.data.frame(),"Dataframes/stats/full metabolite forest.rds")
```
# Abundance Heatmap
We've split them by if they increase or decrease, lets show that and also highlight that they change
```{r}
bch_by_time<-bch_long%>%
  select("BIOCHEMICAL","TIME.POINT.2","ScaledImpData")%>%
  mutate(ScaledImpData=log10(ScaledImpData))%>% # Log transform for plotting
  group_by(TIME.POINT.2,BIOCHEMICAL)%>%
  summarize(mean=mean(ScaledImpData))%>% #Mean of timepoints
  spread(key='TIME.POINT.2',value="mean")%>%
  left_join(.,bch_metadata,by="BIOCHEMICAL")%>%
  filter(BIOCHEMICAL %in% top_bch_forest)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL, levels=rev(top_bch_forest)))%>%
  arrange(BIOCHEMICAL)%>%
  column_to_rownames("BIOCHEMICAL")

pheatmap(bch_by_time[1:4],
         cluster_cols=F,
         cluster_rows=F,
         color=colorRampPalette(c("white","grey","black"),bias=1.5)(50),
         border_color="grey20",
         cellheight = 12,
         cellwidth = 50,
         fontsize = 14,
         filename="Figures/Metabolites/heatmap_sum.pdf",
         width=14,
         height=14
)

bch_all_hmap<-bch_raw%>%
  mutate_if(is.numeric,~log(.))%>%
  left_join(.,bch_metadata%>%select(BIOCHEMICAL,SUPER.PATHWAY,dir),by="BIOCHEMICAL")%>%
  filter(BIOCHEMICAL %in% bch_siglist$BIOCHEMICAL)%>%
  ungroup()%>%
  mutate(pre=rowMeans(.[2:17]))%>%
  mutate(dir=fct_rev(dir))%>%
  arrange(dir,SUPER.PATHWAY,-pre)%>%
  column_to_rownames("BIOCHEMICAL")

pheatmap(bch_all_hmap[1:45],
         cluster_cols=F,
         cluster_rows=F,
         annotation_row=bch_all_hmap%>%select("SUPER.PATHWAY"),
         annotation_col=samplekey%>%select(CLIENT.IDENTIFIER,TIME.POINT.2)%>%column_to_rownames("CLIENT.IDENTIFIER"),
         annotation_colors = anno_color,
         color=colorRampPalette(c("white","grey","black"),bias=1.5)(50),
         border_color="grey20",
         #cellheight = 6.1,
         #cellwidth = 25,
         show_colnames = F,
         annotation_names_col = F,
         annotation_names_row = F,
         fontsize = 10,
         filename="Figures/Supplementals/heatmap_all_signif.pdf",
         width=16,
         height=21
)

```
# Correlate with bacterial families
The big goal with this paper is to look for explanations in changes in metabolites by microbiota and vice versa. A correlation is a nice first step for this
```{r}
cor<-corr.test(bch_for_cor[top_bch_forest],fam_for_cor,method="spearman",adjust="holm")

bch_fam_cor_plot<-left_join(as.data.frame(cor$p.adj)%>%
  rownames_to_column("BIOCHEMICAL")%>%
  gather(key="key",value="p",-BIOCHEMICAL),
  as.data.frame(cor$r)%>%
  rownames_to_column("BIOCHEMICAL")%>%
  gather(key="key",value="r",-BIOCHEMICAL),by=c("BIOCHEMICAL","key"))%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL,levels=rev(top_bch_forest)))%>%
  filter(p<=0.05)%>%
  select(-p)%>%
  complete(BIOCHEMICAL,fill=list(key="drop",r=0))%>%
  spread(key=key,value=r,fill=0)%>%
  column_to_rownames("BIOCHEMICAL")%>%
  select(-drop)

tax_anno<-tax_metadata%>%
  filter(!is.na(family))%>%
  select(family)%>%
  unique()%>%
  remove_rownames()%>%
  filter(family %in% colnames(bch_fam_cor_plot))%>%
  mutate(familyplot=factor(family,levels=names(anno_color$familyplot)))%>%
  column_to_rownames("family")

pheatmap(bch_fam_cor_plot,
         annotation_col = tax_anno,
         cluster_rows=F,
         cluster_cols=T,
         breaks=seq(-1,1,2/60),
         color=c(rev(colorRampPalette(c("grey88","white","dodgerblue3"),bias=2)(30)),colorRampPalette(c("grey88","white", "red3"),bias=2)(30)),
         border_color="grey20",
         cellheight = 12,
         cellwidth = 20,
         fontsize = 14,
         drop_levels = T,
         filename=paste0("Figures/Metabolites/bch_vs_fam_spearman.pdf"),
         annotation_colors = list(familyplot=anno_color$familyplot[names(anno_color$familyplot) %in% unique(tax_anno$familyplot)]),
         width=14,
         height=14)

saveRDS(as.data.frame(cor$p.adj),"Dataframes/Stats/Metabolites vs Family correlation t tests.rds")
```