---
title: "Amino Acid Analysis"
author: "Sam McMillan"
date: "2023-11-30"
output: html_document
---

```{r setup, require=T, include=T}
library(tidyverse)
library(pheatmap)
library(randomForest)
library(ggtext)
library(janitor)
library(writexl)

sessionInfo()
```

#Get Dataframes
```{r}
samplekey<-readRDS("Dataframes/samplekey.rds")
bch_raw<-readRDS("Dataframes/bch_raw.rds")
bch_metadata<-readRDS("Dataframes/bch_metadata.rds")
bch_siglist<-readRDS("Dataframes/Stats/bch_siglist.rds")
fam_siglist<-readRDS("Dataframes/Stats/fam_siglist.rds")
tax_raw<-readRDS("Dataframes/tax_raw.rds")
fam_for_cor<-readRDS("Dataframes/fam_for_cor.rds")
bch_for_cor<-readRDS("Dataframes/bch_for_cor.rds")
tax_metadata<-readRDS("Dataframes/tax_metadata.rds")
pathway_raw<-readRDS("Dataframes/pathway_raw.rds")
pathway_siglist<-readRDS("Dataframes/Stats/pathway_siglist.rds")

source("Scripts/Color codes.R")
```

# Pull Out Amino Acids
We see high amino acids pre FMT, Might be involved in stickland fermentation, lets see
```{r}
AA_raw<-bch_raw%>%
  filter(BIOCHEMICAL %in% c("lysine", "serine", "valine", "alanine",
                          "histidine","glycine","leucine", "proline" ,
                          "arginine","cysteine","tyrosine"  ,
                          "aspartate","glutamate", "glutamine", "threonine",
                          "asparagine","isoleucine", "methionine","tryptophan","phenylalanine",
                          "taurine","ornithine","citrulline","hydroxyproline", #non proteinogenic amino acids
                          "isovalerate (C5)", #oxidative leucine
                          "alpha-hydroxyisocaproate",# reductive leucine
                          "isocaproate (i6:0)", #reductive leucine
                          "5-aminovalerate", # reductive proline
                          "4-hydroxyphenylacetate", #oxidative tyrosine
                          "p-cresol", # oxidative tyrosine
                          "3-(4-hydroxyphenyl)propionate", # reductive tyrosine
                          "indoleacetate", # oxidative Tryptophan
                          "phenylacetate", # oxidative phenylalanine
                          "4-imidazoleacetate", #oxidative histidine
                          "3-phenylpropionate (hydrocinnamate)"##reductive phenylalanine
                          ))

AA_raw_sig<-bch_raw%>%
  left_join(.,bch_metadata,by="BIOCHEMICAL")%>%
  filter(SUPER.PATHWAY %in% c("Amino Acid","Peptide"))%>%
  filter(BIOCHEMICAL %in% bch_siglist$BIOCHEMICAL)
```


# Random Forest
```{r}
AA_long<-AA_raw%>%
  select(BIOCHEMICAL,samplekey$CLIENT.IDENTIFIER)%>%
  gather(key="CLIENT.IDENTIFIER",value="ScaledImpData",-BIOCHEMICAL)%>%
  mutate(CLIENT.IDENTIFIER=factor(CLIENT.IDENTIFIER,levels=levels(samplekey$CLIENT.IDENTIFIER)))%>%
  left_join(.,samplekey,by="CLIENT.IDENTIFIER")

#Random Forest
set.seed(102309)
AA_forest<-AA_raw%>%
  column_to_rownames("BIOCHEMICAL")%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("CLIENT.IDENTIFIER")%>%
  left_join(samplekey,., by="CLIENT.IDENTIFIER")%>%
  select(-CLIENT.IDENTIFIER,-TIME.POINT.2,-SUBJ.ID)%>%
  clean_names()%>%
  randomForest(prepost ~ .,data=., proximity=T, importance=TRUE, ntree=1000)

#Prepare to Plot
AA_forest_plot<-AA_forest$importance%>%
  as.data.frame()%>%
  rownames_to_column("biochem")%>%
  left_join(.,bch_metadata%>%
              select(BIOCHEMICAL,AA_group_detail,dir)%>%
              mutate(biochem=janitor::make_clean_names(BIOCHEMICAL)),by="biochem")%>%
  filter(MeanDecreaseAccuracy>0)%>%
  arrange(-MeanDecreaseAccuracy)%>%
  arrange(dir,MeanDecreaseAccuracy)%>%
  mutate(plotnames=str_remove(BIOCHEMICAL,"\\*"))%>%
  mutate(plotnames=if_else(BIOCHEMICAL %in% bch_siglist$BIOCHEMICAL,paste("*", plotnames),plotnames))%>%
  mutate(plotnames=fct_inorder(plotnames))%>%
  mutate(BIOCHEMICAL=fct_inorder(BIOCHEMICAL))%>%
  mutate(AA_group_detail=factor(AA_group_detail,levels=names(anno_color$AA_group)))

#plot
ggplot(data=AA_forest_plot, aes(y=MeanDecreaseAccuracy,x=plotnames, fill=AA_group_detail))+
  geom_point(shape=21, size=2)+
  scale_fill_manual(values=anno_color$AA_group)+
  coord_flip()+
  theme_bw()+
  labs(fill="", y="Mean Decrease Accuracy",x="")+
  theme(axis.text.y=element_text(size=6),
        axis.text.x=element_text(size=6),
        axis.title.y=element_text(size=8),
        legend.position="top",
        legend.text = element_markdown(size=6),
        legend.key.size = unit(c(0.1,0),"in"),
        legend.box.margin=unit(c(0,1.5,-0.14,0),"in"),
        plot.margin = unit(c(0,0.25,0,0),"in"))+
  guides(fill=guide_legend(nrow=1))

ggsave("Figures/Amino Acids/Amino Acid_forest_sig_all_new.pdf", height=3.3, width=3.5, units="in")

##Error Rate
AA_forest

##Check for not significant, print here and mark in illustrator
AA_forest_plot$BIOCHEMICAL[!AA_forest_plot$BIOCHEMICAL %in% bch_siglist$BIOCHEMICAL]

#make list for filtering
top_AA_forest<-as.vector(AA_forest_plot$BIOCHEMICAL)

## for writing
AA_forest_plot%>%group_by(dir)%>%summarize(count=n())

saveRDS(AA_forest$importance,"Dataframes/Stats/Full amino acid forest.rds")

```

# Abundance Heatmap
```{r}
AA_by_time_hm<-AA_long%>%
  select("BIOCHEMICAL","TIME.POINT.2","ScaledImpData")%>%
  mutate(ScaledImpData=log10(ScaledImpData))%>% # Log transform for plotting
  group_by(TIME.POINT.2,BIOCHEMICAL)%>%
  summarize(mean=mean(ScaledImpData))%>% #Mean of timepoints
  spread(key='TIME.POINT.2',value="mean")%>%
  left_join(.,bch_metadata,by="BIOCHEMICAL")%>%
  filter(BIOCHEMICAL %in% top_AA_forest)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL, levels=rev(top_AA_forest)))%>%
  arrange(BIOCHEMICAL)%>%
  column_to_rownames("BIOCHEMICAL")

pheatmap(AA_by_time_hm[1:4],
         cluster_cols=F,
         cluster_rows=F,
         color=colorRampPalette(c("white","grey","black"),bias=1.5)(50),
         border_color="grey20",
         cellheight = 12,
         cellwidth = 50,
         fontsize = 14,
         filename="Figures/Amino Acids/Amino_acids_heatmap_by_time.pdf",
         width=14,
         height=14
)

```
# family Plots
```{r}
pathway_aa_spec<-pathway_raw%>%
  filter(str_detect(pathway,"Amino-acid") & str_detect(pathway,"\\|"))%>%
  separate(pathway, into=c("pathway","species"),sep="\\|",fill="right")%>%
  #filter(pathway %in% pathway_siglist$pathway)%>%
  separate(pathway,into=c("type","amino","note"),sep=";|\\.",extra="merge")%>%
  separate(amino,into=c("amino","bolism"),sep="_",extra="merge")%>%
  mutate(species=str_remove(species,"\\|"))%>%
  separate(species,into=c("genus","species"),sep=".s__",fill="right")%>%
  mutate(genus=str_remove(genus,"g\\_\\_"))%>%
  left_join(.,tax_metadata%>%select(phylum,family,genus)%>%unique(),by="genus")%>%
  select(amino,bolism,"phylum","family","species",everything())%>%
  mutate(phylum=case_when(!is.na(phylum)~phylum, 
                          genus %in% c("Tyzzerella","Lachnospiraceae_unclassified","Coprococcus","Lachnoclostridium","Subdoligranulum","Ruminococcus","Firmicutes_unclassified","Clostridiales_unclassified")~"Firmicutes", 
                          genus=="Methanobrevibacter" ~ "Euryarchaeota",
                          genus %in% c("Serratia","Vibrio")~"Proteobacteria"))%>%
  mutate(family=case_when(!is.na(family)~family, 
                          genus %in% c("Tyzzerella","Lachnospiraceae_unclassified","Coprococcus","Lachnoclostridium")~"Lachnospiraceae",
                          genus=="Methanobrevibacter" ~ "Methanobacteriaceae",
                          genus %in% c("Subdoligranulum","Ruminococcus")~"Oscillospiraceae",
                          genus=="Serratia"~"Yersiniaceae",
                          genus=="Vibrio"~"Vibrionaceae"))%>%
  select(amino,type,family,samplekey$CLIENT.IDENTIFIER)%>%
  mutate(amino=str_replace(amino,"\\.",""))%>%
  gather(key="CLIENT.IDENTIFIER",value="CPM",-amino,-family,-type)%>%
  group_by(family)%>% #This is to clean up legend
  mutate(max=max(CPM))%>%
  group_by(amino)%>%
  mutate(test=max/max(CPM))%>%
  mutate(family=if_else(family %in% (filter(.,test>0.1))$family,family,"Other"))%>% #has to be at least 10% max to get color
  mutate(CLIENT.IDENTIFIER=factor(CLIENT.IDENTIFIER,levels=unique(samplekey$CLIENT.IDENTIFIER)))%>%
  mutate(family=if_else(is.na(family),"Other",family))%>%
  mutate(family=factor(family,levels=names(anno_color$familyplot)))%>%
  droplevels()

amino_names<-c("glycine"="Glycine",
               "D-alanine"="D-alanine",
               "L-arginine"="Arginine",
               "L-asparagine"="Asparagine",
               "L-cysteine"="Cysteine",
               "L-glutamate"="Glutamate",
               "L-histidine"="Histidine",
               "L-homocysteine"="Homocysteine",
               "L-isoleucine"="Isoleucine",
               "L-leucine"="Leucine",
               "L-lysine"="Lysine",
               "L-methionine"="Methionine",
               "L-phenylalanine"="Phenylalanine",
               "L-proline"="Proline",
               "L-serine"="Serine",
               "L-threonine"="Threonine",
               "L-tryptophan"="Tryptophan",
               "L-tyrosine"="Tyrosine",
               "S-adenosyl-L-methionine"="S-adenosyl-L-methionine"
               )

ggplot(pathway_aa_spec%>%
         mutate(amino=factor(amino,levels=names(amino_names)))%>%
         filter(type=="Amino-acid_biosynthesis")
       ,aes(x=CLIENT.IDENTIFIER,y=CPM, fill=family))+
  geom_bar(stat="identity")+
  theme_bw()+
  facet_wrap(~amino,ncol=5,scales="free_y",labeller = as_labeller(amino_names))+
  theme(axis.text.x=element_text(angle=-90),
        text=element_text(size=8),
        legend.position="right",
        strip.text=element_text(size=8),
        strip.background = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x= element_blank(),
        axis.text.x.bottom = element_blank(),
        legend.box.margin=unit(c(0,0,0,0),"in"),
        legend.text=element_text(size=8))+
  scale_fill_manual(values = anno_color$familyplot)+
  guides(fill=guide_legend(ncol=1))+
  labs(x="",y="Pathway Abundance (CPM)",fill="")

ggsave("Figures/Amino Acids/amino acid biosynthesis_familyplot.pdf",height=5,width=8.5,units="in")
```

# Correlate with bacterial families with repeated measures
```{r}
source("Scripts/rm_corr_all.R")

rmcor_AA_fam<-rmcorr_test(samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
                           "SUBJ.ID",
                           fam_for_cor[levels(fam_siglist$clade_name)],
                           bch_for_cor[top_AA_forest],
                          "BH")

AA_fam_cor_plot<-rmcor_AA_fam%>%
  dplyr::rename("BIOCHEMICAL"=df2,"clade_name"=df1)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL,levels=rev(as.vector(top_AA_forest))))%>%
  filter(clade_name %in% (filter(.,p.adj<=0.05))$clade_name)%>%
  select(-p,-p.adj)%>%
  group_by(clade_name)%>%
  complete(BIOCHEMICAL,fill=list(clade_name="drop",r=0))%>%
   mutate(clade_name=factor(clade_name,levels=names(anno_color$familyplot)))%>%
  spread(key=clade_name,value=r,fill=0)%>%
  column_to_rownames("BIOCHEMICAL")

tax_anno<-tax_metadata%>%
  filter(!is.na(family))%>%
  select(family)%>%
  unique()%>%
  remove_rownames()%>%
  filter(family %in% colnames(AA_fam_cor_plot))%>%
  mutate(familyplot=factor(family,levels=names(anno_color$familyplot)))%>%
  column_to_rownames("family")

AA_fam_cor_p<-rmcor_AA_fam%>%
  dplyr::rename("BIOCHEMICAL"=df2,"clade_name"=df1)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL,levels=rev(as.vector(top_AA_forest))))%>%
  filter(clade_name %in% (filter(.,p.adj<=0.05))$clade_name)%>%
  mutate(ast=symnum(p.adj,cutpoints=c(0,0.05,1),symbols=c("*","")))%>%
  select(-r,-p,-p.adj)%>%
  group_by(clade_name)%>%
  mutate(clade_name=factor(clade_name,levels=names(anno_color$familyplot)))%>%
  complete(BIOCHEMICAL,fill=list(clade_name="drop",ast=""))%>%
  spread(key=clade_name,value=ast,fill="")%>%
  column_to_rownames("BIOCHEMICAL")

pheatmap(AA_fam_cor_plot,
         annotation_col = tax_anno,
         display_numbers=AA_fam_cor_p,
         number_color="white",
         cluster_rows=F,
         cluster_cols=T,
         clustering_distance_cols = "correlation",
         breaks=seq(-1,1,2/60),
         color=c(rev(colorRampPalette(c("grey88","dodgerblue","dodgerblue3"),bias=1)(30)),colorRampPalette(c("grey88","red", "red3"),bias=1)(30)),
         border_color="grey20",
         cellheight = 12,
         cellwidth = 20,
         fontsize = 14,
         drop_levels = T,
         filename=paste0("Figures/Amino Acids/AA_vs_fam_rmcor_BH_all_new.pdf"),
         annotation_colors = list(familyplot=anno_color$familyplot[names(anno_color$familyplot) %in% unique(tax_anno$familyplot)]),
         width=20,
         height=14)

saveRDS(rmcor_AA_fam,"Dataframes/Stats/AA vs Family correlation rmcor.rds")
```

# Random Forest all significant amino acids
```{r}
AA_long_sig<-AA_raw_sig%>%
  select(BIOCHEMICAL,samplekey$CLIENT.IDENTIFIER)%>%
  gather(key="CLIENT.IDENTIFIER",value="ScaledImpData",-BIOCHEMICAL)%>%
  mutate(CLIENT.IDENTIFIER=factor(CLIENT.IDENTIFIER,levels=levels(samplekey$CLIENT.IDENTIFIER)))%>%
  left_join(.,samplekey,by="CLIENT.IDENTIFIER")

#Random Forest
set.seed(102309)
AA_forest_sig<-AA_raw_sig%>%
  column_to_rownames("BIOCHEMICAL")%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("CLIENT.IDENTIFIER")%>%
  left_join(samplekey,., by="CLIENT.IDENTIFIER")%>%
  select(-CLIENT.IDENTIFIER,-TIME.POINT.2,-SUBJ.ID)%>%
  clean_names()%>%
  randomForest(prepost ~ .,data=., proximity=T, importance=TRUE, ntree=1000)

#Prepare to Plot
AA_forest_plot_sig<-AA_forest_sig$importance%>%
  as.data.frame()%>%
  rownames_to_column("biochem")%>%
  left_join(.,bch_metadata%>%
              select(BIOCHEMICAL,AA_group,dir,SUB.PATHWAY)%>%
              mutate(biochem=janitor::make_clean_names(BIOCHEMICAL)),by="biochem")%>%
  filter(MeanDecreaseAccuracy>0)%>%
  arrange(-MeanDecreaseAccuracy)%>%
  arrange(dir,MeanDecreaseAccuracy)%>%
  mutate(plotnames=str_remove(BIOCHEMICAL,"\\*"))%>%
  mutate(plotnames=fct_inorder(plotnames))%>%
  mutate(BIOCHEMICAL=fct_inorder(BIOCHEMICAL))%>%
  mutate(color=if_else(SUB.PATHWAY=="Dipeptide","Dipeptide","Amino Acid"))

#plot
ggplot(data=AA_forest_plot_sig, aes(y=MeanDecreaseAccuracy,x=plotnames, fill=color))+
  geom_point(shape=21, size=2)+
  scale_fill_manual(values=c("deepskyblue2","firebrick4"))+
  coord_flip()+
  theme_bw()+
  labs(fill="", y="Mean Decrease Accuracy",x="")+
  theme(axis.text.y=element_text(size=6),
        axis.text.x=element_text(size=6),
        axis.title.y=element_text(size=8),
        legend.position="top",
        legend.text = element_markdown(size=6),
        legend.key.size = unit(c(0.1,0),"in"),
        legend.box.margin=unit(c(0,1.5,-0.14,0),"in"),
        plot.margin = unit(c(0,0.25,0,0),"in"))+
  guides(fill=guide_legend(ncol=3))
#3.3
ggsave("Figures/Supplementals/Amino Acid_forest_sig_all.pdf", height=4.6, width=3.5, units="in")

##Error Rate
AA_forest_sig

##Check for not significant, print here and mark in illustrator
AA_forest_plot_sig$BIOCHEMICAL[!AA_forest_plot$BIOCHEMICAL %in% bch_siglist$BIOCHEMICAL]

#make list for filtering
top_AA_forest_sig<-as.vector(AA_forest_plot_sig$BIOCHEMICAL)

## for writing
AA_forest_plot_sig%>%group_by(dir)%>%summarize(count=n())

```

# Abundance Heatmap for significant
```{r}
AA_by_time_hm_sig<-AA_long_sig%>%
  select("BIOCHEMICAL","TIME.POINT.2","ScaledImpData")%>%
  mutate(ScaledImpData=log10(ScaledImpData))%>% # Log transform for plotting
  group_by(TIME.POINT.2,BIOCHEMICAL)%>%
  summarize(mean=mean(ScaledImpData))%>% #Mean of timepoints
  spread(key='TIME.POINT.2',value="mean")%>%
  left_join(.,bch_metadata,by="BIOCHEMICAL")%>%
  filter(BIOCHEMICAL %in% top_AA_forest_sig)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL, levels=rev(top_AA_forest_sig)))%>%
  arrange(BIOCHEMICAL)%>%
  column_to_rownames("BIOCHEMICAL")

pheatmap(AA_by_time_hm_sig[1:4],
         cluster_cols=F,
         cluster_rows=F,
         color=colorRampPalette(c("white","grey","black"),bias=1.5)(50),
         border_color="grey20",
         cellheight = 12,
         cellwidth = 50,
         fontsize = 14,
         filename="Figures/Supplementals/Amino_acids_heatmap_by_time_sig_all.pdf",
         width=14,
         height=14
)

```

# Correlate with bacterial families with repeated measures for significant amino acids
```{r}
source("Scripts/rm_corr_all.R")

rmcor_AA_fam_sig<-rmcorr_test(samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
                           "SUBJ.ID",
                           fam_for_cor[levels(fam_siglist$clade_name)],
                           bch_for_cor[top_AA_forest_sig],
                          "BH")

AA_fam_cor_plot_sig<-rmcor_AA_fam_sig%>%
  dplyr::rename("BIOCHEMICAL"=df2,"clade_name"=df1)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL,levels=rev(as.vector(top_AA_forest_sig))))%>%
  filter(clade_name %in% (filter(.,p.adj<=0.05))$clade_name)%>%
  select(-p,-p.adj)%>%
  group_by(clade_name)%>%
  complete(BIOCHEMICAL,fill=list(clade_name="drop",r=0))%>%
   mutate(clade_name=factor(clade_name,levels=names(anno_color$familyplot)))%>%
  spread(key=clade_name,value=r,fill=0)%>%
  column_to_rownames("BIOCHEMICAL")

tax_anno_sig<-tax_metadata%>%
  filter(!is.na(family))%>%
  select(family)%>%
  unique()%>%
  remove_rownames()%>%
  filter(family %in% colnames(AA_fam_cor_plot_sig))%>%
  mutate(familyplot=factor(family,levels=names(anno_color$familyplot)))%>%
  column_to_rownames("family")

AA_fam_cor_p_sig<-rmcor_AA_fam_sig%>%
  dplyr::rename("BIOCHEMICAL"=df2,"clade_name"=df1)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL,levels=rev(as.vector(top_AA_forest_sig))))%>%
  filter(clade_name %in% (filter(.,p.adj<=0.05))$clade_name)%>%
  mutate(ast=symnum(p.adj,cutpoints=c(0,0.05,1),symbols=c("*","")))%>%
  select(-r,-p,-p.adj)%>%
  group_by(clade_name)%>%
  mutate(clade_name=factor(clade_name,levels=names(anno_color$familyplot)))%>%
  complete(BIOCHEMICAL,fill=list(clade_name="drop",ast=""))%>%
  spread(key=clade_name,value=ast,fill="")%>%
  column_to_rownames("BIOCHEMICAL")

pheatmap(AA_fam_cor_plot_sig,
         annotation_col = tax_anno_sig,
         display_numbers=AA_fam_cor_p_sig,
         number_color="white",
         cluster_rows=F,
         cluster_cols=T,
         clustering_distance_cols = "correlation",
         breaks=seq(-1,1,2/60),
         color=c(rev(colorRampPalette(c("grey88","dodgerblue","dodgerblue3"),bias=1)(30)),colorRampPalette(c("grey88","red", "red3"),bias=1)(30)),
         border_color="grey20",
         cellheight = 12,
         cellwidth = 20,
         fontsize = 14,
         drop_levels = T,
         filename=paste0("Figures/Supplementals/AA_vs_fam_rmcor_sig.pdf"),
         annotation_colors = list(familyplot=anno_color$familyplot[names(anno_color$familyplot) %in% unique(tax_anno_sig$familyplot)]),
         width=20,
         height=14)

```

#Data export for Mendeley
```{r data export}
write_xlsx(list("4A"=AA_forest_plot,
                "4A linear mixed model"=bch_siglist,
                "4B"=AA_by_time_hm,
                "4C"=AA_fam_cor_plot,
                "4C linear model"=rmcor_AA_fam,
                "4D"=pathway_aa_spec),
           "Excel Exports/Figure 4.xlsx")

write_xlsx(list("S2A"=AA_forest_plot_sig,
                "S2B"=AA_by_time_hm_sig,
                "S2C"=AA_fam_cor_plot_sig,
                "S2C linear model"=rmcor_AA_fam_sig),
           "Excel Exports/Figure S2.xlsx")

```


