---
title: "Carnitine Metabolism Analysis"
author: "Sam McMillan"
date: "2023-11-30"
output: html_document
---

```{r setup, require=T, include=T}
library(tidyverse)
library(ggpubr)
library(pheatmap)
library(rstatix)

sessionInfo()
```

#Get Dataframes
```{r}
samplekey<-readRDS("Dataframes/samplekey.rds")
pathway_raw<-readRDS("Dataframes/pathway_raw.rds")
bch_siglist<-readRDS("Dataframes/Stats/bch_siglist.rds")
bch_metadata<-readRDS("Dataframes/bch_metadata.rds")
bch_for_cor<-readRDS("Dataframes/bch_for_cor.rds")
tax_metadata<-readRDS("Dataframes/tax_metadata.rds")
top_bch_forest<-readRDS("Dataframes/top_bch_forest.rds")

source("Scripts/Color codes.R")
```

# Pull Out Carnitine Metabolism
isolate carnitine metabolism to see whats going on with those high carnitines in the metabolites dataset
```{r}
carn<-pathway_raw%>%
  filter(str_detect(pathway,"carnitine|Carnitine"))%>%
  separate(pathway,into=c("UniRef","taxa"),sep = "\\|",fill="right")%>%
  filter(!is.na(taxa))%>%
  separate(taxa,into=c("genus","species"),sep = "\\.",fill="right")%>%
  mutate(genus=str_remove(genus,"g\\_\\_"))%>%
  left_join(.,tax_metadata%>%select(family,phylum,genus)%>%unique(),by="genus")%>%
  select("UniRef","phylum","genus","species",everything())

carn_long<-carn%>%
  gather(key="CLIENT.IDENTIFIER",value="CPM",-c("UniRef","phylum","family","genus","species"))%>%
  mutate(CLIENT.IDENTIFIER=factor(CLIENT.IDENTIFIER,levels=levels(samplekey$CLIENT.IDENTIFIER)))%>%
  left_join(.,samplekey,by="CLIENT.IDENTIFIER")

saveRDS(carn,"Dataframes/carnitine metabolism.rds")
```

# Boxplot
plot all the carnitine metabolism for each timepoint and see if it signficantly changes
```{r}
carn_by_time<-carn_long%>%
  select(CLIENT.IDENTIFIER,CPM,TIME.POINT.2,SUBJ.ID)%>%
  group_by(CLIENT.IDENTIFIER,TIME.POINT.2,SUBJ.ID)%>%
  summarize(sum=sum(CPM))

carn_boxplot_stats<-carn_by_time%>%
  ungroup()%>%
  select(sum,TIME.POINT.2,SUBJ.ID)%>%
  complete(TIME.POINT.2, SUBJ.ID, fill=list(sum=NA))%>%
  arrange(SUBJ.ID)%>%
  pairwise_wilcox_test(sum ~ TIME.POINT.2,
                       p.adjust.method="holm",
                       ref.group = "Pre",
                       paired=T)%>%
  add_xy_position(x="TIME.POINT.2",step.increase = 0.08)

ggplot(carn_by_time, aes(x=TIME.POINT.2,y=sum,group=TIME.POINT.2))+
  geom_dotplot(aes(fill=TIME.POINT.2),binaxis="y",stackdir="center",binwidth=100/25,dotsize=1,stroke=0.5)+
  geom_boxplot(alpha=0,linewidth=0.2)+
  theme_bw()+
  theme(text=element_text(size=8),
        legend.position="top",
        legend.text = element_text(size=7),
        legend.key.size = unit(0.1 ,'in'),
        legend.box.margin = unit(c(0,.5,0,0),"in"),
        axis.title.y = element_text(size=6),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x= element_blank())+
  labs(y = "Carnitine Metabolism Pathway Abundance (CPM)",
       x= "",
       fill="")+
  guides(fill=guide_legend(nrow=1))+
  scale_y_continuous(limits=c(0,100),expand=c(0,0))+
  scale_fill_manual(values=anno_color$TIME.POINT.2)+
 stat_pvalue_manual(filter(carn_boxplot_stats,p.adj<0.05), 
                     label = "p.adj.signif",
                     tip.length=0.01,
                     step.increase=0,
                     inherit.aes=F)

ggsave("Figures/Carnitine/carn_boxplot.pdf",height=2.5,width=3, units="in")

saveRDS(carn_boxplot_stats,"Dataframes/Stats/Carnitine metabolism_wilcox_signed_rank_holm.rds")
```

#Species Barplot
Check which species are encoding carnitine metabolic pathways
```{r}
carn_spec_plot<-carn_long%>%
  mutate(species=str_remove(species,"s__"))%>%
  mutate(species=str_replace_na(species,"Other"))%>%
  mutate(species=str_replace_all(species,"_"," "))%>%
  arrange(-CPM)%>%
  mutate(species=fct_inorder(species))%>%
  mutate(species=fct_relevel(species,"Other",after=Inf))%>%
  droplevels()

ggplot(carn_spec_plot,aes(x=CLIENT.IDENTIFIER,y=CPM, fill=species))+
  geom_bar(stat="identity")+
  theme_bw()+
  theme(axis.text.x=element_blank(),
        text=element_text(size=8),
        legend.text = element_text(size=7),
        legend.key.size = unit(0.1 ,'in'),
        axis.title.y = element_text(size=6),
        legend.box.margin = unit(c(0,0,0,-0.15),"in"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x= element_blank())+
  scale_fill_manual(values = anno_color$speciesplot)+
  scale_y_continuous(limits=c(0,70),expand=c(0,0))+
  labs(x="",y="Carnitine Metabolism Pathway Abundance (CPM)",fill="Species")
ggsave("Figures/Carnitine/carn_family_plot.pdf",height=2,width=3.5, units="in")
```
its all enterobacteriaceae, neat

#Carnitine Metabolism Correlation with repeated measures
are some bacterial species present alongside certain species of carnitines?
```{r}
carn_for_cor<-carn%>%
  mutate(species=str_remove(species,"s__"))%>%
  mutate(species=str_replace_na(species,"Unclassified"))%>%
  mutate(species=str_replace_all(species,"_"," "))%>%
  select(species,levels(samplekey$CLIENT.IDENTIFIER))%>%
  column_to_rownames("species")%>%
  t()

sig_carn<-bch_metadata%>%
  filter(SUB.PATHWAY=="Fatty Acid Metabolism(Acyl Carnitine)"| str_detect(BIOCHEMICAL,"carnitine|Carnitine"))%>%
  filter(BIOCHEMICAL %in% bch_siglist$BIOCHEMICAL)

source("Scripts/rm_corr_all.R")
rmcor_carn_spec<-rmcorr_test(samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
                           "SUBJ.ID",
                           carn_for_cor,
                          bch_for_cor[sig_carn$BIOCHEMICAL],
                          "BH")

carn_spec_cor_plot<-rmcor_carn_spec%>%
  dplyr::rename("BIOCHEMICAL"=df2,"clade_name"=df1)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL,levels=as.vector(sig_carn$BIOCHEMICAL)))%>%
  mutate(clade_name=factor(str_replace(clade_name,"Unclassified","Other"), levels=rev(levels(carn_spec_plot$species))))%>%
  filter(clade_name %in% (filter(.,p.adj<=0.05))$clade_name)%>%
  select(-p,-p.adj)%>%
  complete(BIOCHEMICAL,fill=list(clade_name="drop",r=0))%>%
  spread(key=clade_name,value=r,fill=0)%>%
  mutate(BIOCHEMICAL=str_remove_all(BIOCHEMICAL,"\\*"))%>%
  column_to_rownames("BIOCHEMICAL")

carn_spec_cor_p<-rmcor_carn_spec%>%
  dplyr::rename("BIOCHEMICAL"=df2,"clade_name"=df1)%>%
  mutate(BIOCHEMICAL=factor(BIOCHEMICAL,levels=as.vector(sig_carn$BIOCHEMICAL)))%>%
  mutate(clade_name=factor(str_replace(clade_name,"Unclassified","Other"), levels=rev(levels(carn_spec_plot$species))))%>%
  filter(clade_name %in% (filter(.,p.adj<=0.05))$clade_name)%>%
  mutate(ast=symnum(p.adj,cutpoints=c(0,0.05,1),symbols=c("*","")))%>%
  select(-r,-p,-p.adj)%>%
  complete(BIOCHEMICAL,fill=list(clade_name="drop",r=0))%>%
  spread(key=clade_name,value=ast,fill=0)%>%
  mutate(BIOCHEMICAL=str_remove_all(BIOCHEMICAL,"\\*"))%>%
  column_to_rownames("BIOCHEMICAL")

source("Scripts/pheatmap_left.R")

pheatmap(carn_spec_cor_plot,
         display_numbers=carn_spec_cor_p,
         number_color="white",
         cluster_rows=F,
         cluster_cols=F,
         fontsize=16,
         annotation_colors=anno_color,
         breaks=seq(-1,1,2/60),
         color=c(rev(colorRampPalette(c("grey88","dodgerblue","dodgerblue3"),bias=1)(30)),colorRampPalette(c("grey88","red", "red3"),bias=1)(30)),
         angle_col=270,
         width=6,
         height=5,
         filename=paste0("Figures/Carnitine/bch_vs_carn_rmcor_BH_all.pdf"))

##Reset the pheatmap left overwrite
detach("package:pheatmap", unload = TRUE)
library(pheatmap)

saveRDS(as.data.frame(carn_cor$p.adj),"Dataframes/stats/Carn metabolism vs Carn correlation t test holm.rds")
```


# For Writing
I mention some stuff in the text, these are the summaries
```{r}
# for writing
#total acylcarnitines in metabolomics
nrow(bch_metadata%>%filter(str_detect(BIOCHEMICAL,"carnitine")))-2 #subtract out "carnitine" and "deoxycarnitine" I confirmed succinyl and malonyl are acyl carnitines.
# how many were significant
nrow(bch_siglist%>%select(BIOCHEMICAL)%>%unique()%>%filter(str_detect(BIOCHEMICAL,"carnitine")))
#how many significant are in random forest
nrow(bch_siglist%>%select(BIOCHEMICAL)%>%unique()%>%filter(str_detect(BIOCHEMICAL,"carnitine") & BIOCHEMICAL %in% top_bch_forest))


```

#Data export for Mendeley
```{r data export}
write_xlsx(list("3A"=carn_by_time,
                "3A Wilcoxon signed rank"=carn_boxplot_stats,
                "3B"=carn_spec_plot,
                "3C"=carn_spec_cor_plot,
                "3C linear model"=rmcor_carn_spec),
           "Excel Exports/Figure 3.xlsx")

```

