---
title: "MCBAs"
author: "Sam McMillan"
date: "2023-12-06"
output: html_document
---

# setup
```{r setup, include=T}
library(tidyverse)
library(pheatmap)
library(randomForest)
library(janitor)
library(xlsx)
library(ggtext)
library(vegan)
library(ggpubr)
library(rstatix)
library(writexl)

sessionInfo()
```

# get data
```{r}

samplekey<-readRDS("Dataframes/samplekey.rds")
pathway_raw<-readRDS("Dataframes/pathway_raw.rds")
fam_for_cor<-readRDS("Dataframes/fam_for_cor.rds")
bch_for_cor<-readRDS("Dataframes/bch_for_cor.rds")
tax_metadata<-readRDS("Dataframes/tax_metadata.rds")
tax_raw<-readRDS("Dataframes/tax_raw.rds")
fam_siglist<-readRDS("Dataframes/Stats/fam_siglist.rds")
MCBA_raw<-readRDS("Dataframes/MCBA_raw.rds")
MCBA_metadata<-readRDS("Dataframes/MCBA_metadata.rds")
MCBA_siglist<-readRDS("Dataframes/Stats/MCBA_siglist.rds")
MCBA_for_cor<-readRDS("Dataframes/MCBA_for_cor.rds")
pathway_siglist<-readRDS("Dataframes/Stats/pathway_siglist.rds")

source("Scripts/Color codes.R")
```

# Scale Data
I want to median scale it to be similar to metabolon, but I don't like the minimum imputation so I'll leave them as NA
```{r}
MCBA_median_scaled<-MCBA_raw%>%
  filter(rowSums(.[2:46]==0)<45)%>% #filter out MCBA with no responses
  gather(key="CLIENT.IDENTIFIER",value="value",-BA)%>%
  mutate(CLIENT.IDENTIFIER=factor(CLIENT.IDENTIFIER,levels=levels(samplekey$CLIENT.IDENTIFIER)))%>%
  group_by(BA)%>%
  mutate(BA=fct_inorder(BA))%>%
  filter(value!=0)%>%
  mutate(value=value/median(value))%>%
  as.data.frame()%>%
  complete(CLIENT.IDENTIFIER, BA, fill=list(value=NA))%>%
  spread(key="CLIENT.IDENTIFIER",value="value")
```

# Random Forest
```{r}
MCBA_dir<-MCBA_siglist%>%
  mutate(dir=factor(if_else(coef>0,"Up","Down"),levels=c("Up","Down")))%>%
  select(BA,dir)

#Random Forest
set.seed(102309)
MCBA_forest<-MCBA_raw%>%
  column_to_rownames("BA")%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("CLIENT.IDENTIFIER")%>%
  left_join(samplekey,., by="CLIENT.IDENTIFIER")%>%
  select(-CLIENT.IDENTIFIER,-TIME.POINT.2,-SUBJ.ID)%>%
  clean_names()%>%
  randomForest(prepost ~ .,data=., proximity=T, importance=TRUE, ntree=1000)

#Prepare to Plot
MCBA_forest_plot<-MCBA_forest$importance%>%
  as.data.frame()%>%
  rownames_to_column("ba")%>%
  left_join(.,MCBA_metadata%>%mutate(ba=janitor::make_clean_names(BA)),by="ba")%>%
  left_join(MCBA_dir,by="BA")%>%
  arrange(-MeanDecreaseAccuracy)%>%
  group_by(dir)%>%
  arrange(dir,MeanDecreaseAccuracy)%>%
  filter(BA %in% MCBA_siglist$BA)%>%
  filter(MeanDecreaseAccuracy>0)%>%
  unique()%>%
  mutate(plotnames=str_remove(BA,"\\*"))%>%
  mutate(plotnames=if_else(BA %in% MCBA_siglist$BA,paste("*", plotnames),plotnames))%>%
  mutate(plotnames=fct_inorder(plotnames))%>%
  mutate(BA=fct_inorder(BA))

#plot
ggplot(data=MCBA_forest_plot, aes(y=MeanDecreaseAccuracy,x=BA,fill=Core))+
  geom_point(size=2,shape=21)+
  scale_fill_manual(values=anno_color$Core)+
  coord_flip()+
  theme_bw()+
  labs(fill="",shape="", y="Mean Decrease Accuracy        ",x="")+
  theme(axis.text.y=element_text(size=6),
        axis.text.x=element_text(size=6),
        axis.title.y=element_text(size=8),
        legend.position="top",
        legend.text = element_text(size=6),
        legend.key.size = unit(c(0),"in"),
        legend.box.margin=unit(c(0,0.5,-.05,0),"in"),
        plot.margin = unit(c(0,0.25,0,0),"in"))

ggsave("Figures/MCBA/MCBA_forest_sig.pdf", height=4.2, width=2.5, units="in")

# For writing
#MCBAs with no signal
MCBA_raw%>%filter(rowSums(.[2:46]==0)==45)%>%select(BA)

#number of singificant
length(unique(MCBA_siglist$BA))

#Significant MCBAs not in RFA
MCBA_siglist%>%filter(!BA %in% MCBA_forest_plot$BA)

#number of important
nrow(MCBA_forest$importance%>%
  as.data.frame()%>%
  rownames_to_column("ba")%>%
  left_join(.,MCBA_metadata%>%mutate(ba=janitor::make_clean_names(BA)),by="ba")%>%
  left_join(MCBA_dir,by="BA")%>%
  arrange(-MeanDecreaseAccuracy)%>%
  group_by(dir)%>%
  unique()%>%
  mutate(count=1:n())%>%
  arrange(dir,MeanDecreaseAccuracy)%>%
  filter(MeanDecreaseAccuracy>0))

#number up and down total
MCBA_forest_plot%>%
  group_by(dir)%>%
  summarize(count=n())
#number up and down split by core
MCBA_forest_plot%>%
  group_by(dir,Core)%>%
  summarize(count=n())%>%
  spread(dir,count)
#number up and down split by amino acid
MCBA_forest_plot%>%
  group_by(dir,Amino_Acid)%>%
  summarize(count=n())%>%
  spread(dir,count)

saveRDS(as.vector(MCBA_forest_plot$BA),"Dataframes/MCBA order.rds")
saveRDS(MCBA_forest$importance,"Dataframes/Stats/Full MCBA forest.rds")
```


# heatmap
```{r}
# Abundance Heatmap
MCBA_by_time<-MCBA_median_scaled%>%
  gather(key="CLIENT.IDENTIFIER",value="value",-BA)%>%
  mutate(CLIENT.IDENTIFIER=factor(CLIENT.IDENTIFIER,levels=levels(samplekey$CLIENT.IDENTIFIER)))%>%
  left_join(.,samplekey,by="CLIENT.IDENTIFIER")%>%
  select(BA,TIME.POINT.2,value)%>%
  group_by(TIME.POINT.2,BA)%>%
  replace_na(list(value=0))%>%
  summarize(mean=mean(value))%>% #Mean of timepoints
  mutate(mean=if_else(mean==0,NA,mean))%>%
  spread(key='TIME.POINT.2',value="mean")%>%
  filter(BA %in% MCBA_forest_plot$BA)%>%
  mutate(BA=factor(BA, levels=rev(MCBA_forest_plot$BA)))%>%
  arrange(BA)%>%
  mutate_if(is.numeric,log10)%>%
  left_join(.,MCBA_metadata,by="BA")%>%
  column_to_rownames("BA")

pheatmap(MCBA_by_time[1:4],
         cluster_cols=F,
         cluster_rows=F,
         color=colorRampPalette(c("white","grey","black"),bias=1.5)(50),
         border_color="grey20",
         cellheight = 12,
         na_col="white",
         cellwidth = 50,
         fontsize = 14,
         filename="Figures/MCBA/MCBA_heatmap_RF order.pdf",
         width=14,
         height=14
)

# All MCBA Heatmap sort
MCBA_hm_All<-MCBA_median_scaled%>%
  mutate_if(is.numeric,log10)%>%
  left_join(.,MCBA_metadata,by="BA")%>%
  #filter(BA %in% MCBA_siglist$BA)%>%
  column_to_rownames("BA")%>%
  mutate_if(is.numeric,~replace(., is.na(.), -5)) #Fill with very small 10e-5 for clustering, also makes non-zero visible

pheatmap(MCBA_hm_All%>%arrange(Core,Amino_Acid)%>%select(1:45),
         annotation_row=MCBA_hm_All[c(46,47)],
         annotation_col=samplekey%>%select(CLIENT.IDENTIFIER,TIME.POINT.2)%>%column_to_rownames("CLIENT.IDENTIFIER"),
         annotation_colors = anno_color,
         cluster_cols=T,
         cluster_rows=T,
         color=colorRampPalette(c("white","grey","black"),bias=1)(50),
         border_color="grey20",
         na_col="white",
         fontsize = 6,
         filename="Figures/MCBA/MCBA_heatmap_all_xy.pdf",
         width=8,
         height=11
)
```
# Correlate with Family Repeated measures
```{r}

source("Scripts/rm_corr_all.R")
rmcor_MCBA_fam<-rmcorr_test(samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
                           "SUBJ.ID",
                           fam_for_cor[levels(fam_siglist$clade_name)],
                           MCBA_for_cor[MCBA_forest_plot$BA],
                          "BH")

MCBA_fam_cor_plot<-rmcor_MCBA_fam%>%
  dplyr::rename("BA"=df2,"clade_name"=df1)%>%
  mutate(BA=factor(BA, levels=rev(MCBA_forest_plot$BA)))%>%
  filter(!is.na(BA))%>%
  filter(clade_name %in% (filter(.,p.adj<=0.05))$clade_name)%>%
  select(-p,-p.adj)%>%
  group_by(clade_name)%>%
  mutate(clade_name=factor(clade_name,levels=names(anno_color$familyplot)))%>%
  complete(BA,fill=list(clade_name="drop",r=0))%>%
  spread(key=clade_name,value=r,fill=0)%>%
  column_to_rownames("BA")

tax_anno<-tax_metadata%>%
  filter(!is.na(family))%>%
  select(family)%>%
  unique()%>%
  remove_rownames()%>%
  filter(family %in% colnames(MCBA_fam_cor_plot))%>%
  mutate(familyplot=factor(family,levels=names(anno_color$familyplot)))%>%
  column_to_rownames("family")

MCBA_fam_cor_p<-rmcor_MCBA_fam%>%
  dplyr::rename("BA"=df2,"clade_name"=df1)%>%
  mutate(BA=factor(BA,levels=rev(MCBA_forest_plot$BA)))%>%
  filter(!is.na(BA))%>%
  filter(clade_name %in% (filter(.,p.adj<=0.05))$clade_name)%>%
  mutate(ast=symnum(p.adj,cutpoints=c(0,0.05,1),symbols=c("*","")))%>%
  select(-r,-p,-p.adj)%>%
  group_by(clade_name)%>%
  mutate(clade_name=factor(clade_name,levels=names(anno_color$familyplot)))%>%
  complete(BA,fill=list(clade_name="drop",r=0))%>%
  spread(key=clade_name,value=ast,fill="")%>%
  column_to_rownames("BA")

pheatmap(MCBA_fam_cor_plot,
         annotation_col = tax_anno,
         display_numbers=MCBA_fam_cor_p,
         number_color="white",
         cluster_rows=F,
         cluster_cols=T,
         clustering_distance_cols = "correlation",
         breaks=seq(-1,1,2/60),
         color=c(rev(colorRampPalette(c("grey88","dodgerblue","dodgerblue3"),bias=1)(30)),colorRampPalette(c("grey88","red", "red3"),bias=1)(30)),
         border_color="grey20",
         cellheight = 12,
         cellwidth = 20,
         fontsize = 14,
         drop_levels = T,
         filename=paste0("Figures/MCBA/MCBA_vs_fam_rmcor_BH_all_new.pdf"),
         annotation_colors = list(familyplot=anno_color$familyplot[names(anno_color$familyplot) %in% unique(tax_anno$familyplot)]),
         width=20,
         height=14)

saveRDS(rmcor_MCBA_fam,"Dataframes/Stats/MCBA vs family rmcor BH.rds")

```


# Boxplot by time (NOT IN USE)
```{r}
MCBA_boxplot<-MCBA_raw%>%
  gather(key="CLIENT.IDENTIFIER",value="value",-BA)%>%
  left_join(.,MCBA_metadata,by="BA")%>%
  left_join(.,samplekey,by="CLIENT.IDENTIFIER")%>%
  mutate()

MCBA_boxplot_stats<-MCBA_boxplot%>%
  ungroup()%>%
  select(value,TIME.POINT.2,SUBJ.ID, BA)%>%
  complete(TIME.POINT.2, SUBJ.ID, BA, fill=list(value=NA))%>%
  left_join(.,MCBA_metadata%>%select(BA,Core,Amino_Acid),by="BA")%>%
  group_by(Core,Amino_Acid)%>%
    pairwise_wilcox_test(value ~ TIME.POINT.2,
                       p.adjust.method="holm",
                       ref.group = "Pre",
                       paired=T)%>%
  add_xy_position(x="Core")%>%
  mutate(sort=paste0(Core,Amino_Acid))%>%
  mutate(y.position=log10(y.position)*1.1)%>%
  mutate(xmin=ifelse(xmin>4,xmin-1,xmin))%>%
  mutate(xmax=ifelse(xmax>4,xmax-1,xmax))

#list of aa significant by linear modelling
target_aa<-c("Tau","Gly","Unconjugated","Ala","Ile/Leu","Val","Phe","Trp","Tyr","Ser" ,"Pro","Glu","Lys")

MCBA_boxplot_sort<-MCBA_boxplot%>%
         filter(Core %in%filter(MCBA_boxplot_stats,p.adj<0.05)$Core &
           Amino_Acid %in% target_aa)%>%
         mutate(Amino_Acid=factor(Amino_Acid,levels=target_aa))

ggplot(data=MCBA_boxplot_sort,
       aes(x=Core,y=value,fill=TIME.POINT.2, group=interaction(Core,TIME.POINT.2)))+
  geom_point(shape=21,position=position_dodge(width=0.75),size=1,stroke=0.1)+
  geom_boxplot(alpha=0,linewidth=0.2)+
  facet_wrap(~Amino_Acid,ncol=3)+
  theme_bw()+
  theme(text=element_text(size=10),
        legend.position="none",
        axis.title.y=element_markdown(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linewidth=0.2),
        panel.grid.minor= element_blank(),
        strip.background=element_blank())+
  labs(y = "MCBA Abundance",
       x= "",
       fill="Time Post FMT")+
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     breaks=c(0,1e2,1e4,1e6,1e8,1e10),
                     expand=c(0.03,0,0.07,0))+
  scale_fill_manual(values=anno_color$TIME.POINT.2)+
  stat_pvalue_manual(filter(MCBA_boxplot_stats,p.adj<0.05)%>%
         mutate(Amino_Acid=factor(Amino_Acid,levels=levels(MCBA_boxplot_sort$Amino_Acid))), 
                     label = "p.adj.signif",
                     size=2,
                     tip.length=0.01,
                     step.increase=0.07,
                     step.group.by="sort",
                     inherit.aes=F)

ggsave("Figures/MCBA/MCBA_Boxplots_by_time_wilcox.pdf",width=8.25, height=6)


##Everything
ggplot(data=MCBA_boxplot,
       aes(x=Core,y=value,fill=TIME.POINT.2, group=interaction(Core,TIME.POINT.2)))+
  geom_point(shape=21,position=position_dodge(width=0.75),size=1,stroke=0.1)+
  geom_boxplot(alpha=0,linewidth=0.2)+
  facet_wrap(~Amino_Acid,ncol=5)+
  theme_bw()+
  theme(text=element_text(size=10),
        legend.position="none",
        axis.title.y=element_markdown(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linewidth=0.2),
        panel.grid.minor= element_blank(),
        axis.text.x=element_text(angle=-90),
        strip.background=element_blank())+
  labs(y = "MCBA Abundance",
       x= "",
       fill="Time Post FMT")+
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     breaks=c(0,1e2,1e4,1e6,1e8,1e10),
                     expand=c(0.03,0,0.07,0))+
  scale_fill_manual(values=anno_color$TIME.POINT.2)+
  stat_pvalue_manual(filter(MCBA_boxplot_stats,p.adj<0.05)%>%
         mutate(Amino_Acid=factor(Amino_Acid,levels=levels(MCBA_boxplot_sort$Amino_Acid))), 
                     label = "p.adj.signif",
                     size=2,
                     tip.length=0.01,
                     step.increase=0.07,
                     step.group.by="sort",
                     inherit.aes=F)

ggsave("Figures/MCBA/MCBA_Boxplots_all_by_time_wilcox.pdf",width=8.25, height=8.5)
```

# Pre/Post Boxplot
```{r}
MCBA_boxplot<-MCBA_raw%>%
  gather(key="CLIENT.IDENTIFIER",value="value",-BA)%>%
  left_join(.,MCBA_metadata,by="BA")%>%
  left_join(.,samplekey,by="CLIENT.IDENTIFIER")


MCBA_siglist_prepost<-read_tsv("Maaslin output/MCBA_raw_prepost/all_results.tsv")%>%
            left_join(.,MCBA_raw%>%select(BA)%>%mutate(feature=janitor::make_clean_names(BA)), by="feature")%>%
  left_join(.,MCBA_metadata,by="BA")

MCBA_boxplot_stats<-MCBA_boxplot%>%
  ungroup()%>%
  select(value,prepost,SUBJ.ID, BA)%>%
  complete(prepost, SUBJ.ID, BA, fill=list(value=NA))%>%
  left_join(.,MCBA_metadata%>%select(BA,Core,Amino_Acid),by="BA")%>%
  group_by(Core,Amino_Acid)%>%
  pairwise_t_test(value ~ prepost,  ##I'm using this for the convenient structure, will overwrite with linear modeling results
                       p.adjust.method="holm",
                       ref.group = "Pre")%>%
  add_xy_position(x="Core")%>%
  select(-c(p,p.signif,p.adj,p.adj.signif))%>%
  left_join(.,MCBA_siglist_prepost%>%
              select(BA,pval,qval,Core,Amino_Acid),
            by=c("Core","Amino_Acid"))%>%
  ungroup()%>%
  mutate(q.adj.signif=symnum(qval,
                             cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), 
                             symbols = c("****", "***", "**", "*", "")))%>%
  mutate(sort=paste0(Core,Amino_Acid))%>%
  mutate(y.position=log10(y.position)*1.1)

#sigAA
(MCBA_siglist%>%left_join(.,MCBA_metadata,by="BA")%>%ungroup()%>%select(Amino_Acid)%>%unique())$Amino_Acid

#list of aa significant by linear modelling
aa_names<-c("Tau"="Taurine","Gly"="Glycine","Ala"="Alanine","Glu"="Glutamate","His"="Histidine","Ile_Leu"="Isoleucine/Leucine","Lys"="Lysine","Phe"="Phenylalanine","Ser"="Serine","Trp"="Tryptophan","Tyr"="Tyrosine","Val"="Valine")

MCBA_boxplot_sort<-MCBA_boxplot%>%
         filter(Amino_Acid %in% names(aa_names))%>%
         mutate(Amino_Acid=factor(Amino_Acid,levels=names(aa_names)))

ggplot(data=MCBA_boxplot_sort,aes(x=Core,y=value,fill=prepost, group=interaction(Core,prepost)))+
  geom_jitter(shape=21,
              position=position_jitterdodge(dodge.width=0.75,jitter.width=0.2),
              size=0.8,stroke=0.1)+
  geom_boxplot(alpha=0,linewidth=0.2)+
  facet_wrap(~Amino_Acid,ncol=3,labeller=labeller(Amino_Acid=aa_names))+
  theme_bw()+
  theme(text=element_text(size=10),
        legend.position="none",
        axis.title.y=element_markdown(),
        axis.text.x=element_text(angle=-90),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linewidth=0.2),
        panel.grid.minor= element_blank(),
        strip.background=element_blank())+
  labs(y = "MCBA Abundance",
       x= "",
       fill="Time Post FMT")+
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     breaks=c(0,1e2,1e4,1e6,1e8,1e10),
                     expand=c(0.03,0,0.1,0))+
  scale_fill_manual(values=anno_color$TIME.POINT.2)+
  stat_pvalue_manual(filter(MCBA_boxplot_stats,qval<0.05)%>%
                      filter(Amino_Acid %in% names(aa_names))%>%
                       droplevels()%>%
                      mutate(x=as.numeric(Core))%>%
                      mutate(xmin=x-0.2)%>%
                      mutate(xmax=x+0.2)%>%
                      mutate(Amino_Acid=factor(Amino_Acid,levels=levels(MCBA_boxplot_sort$Amino_Acid))), 
                     label = "q.adj.signif",
                     size=2,
                     tip.length=0.01,
                     step.increase=0.07,
                     step.group.by="sort",
                     inherit.aes=F)

ggsave("Figures/MCBA/MCBA_Boxplots.pdf",width=8.25, height=4)


#Deconjugated 
(MCBA_siglist%>%left_join(.,MCBA_metadata,by="BA")%>%filter(Amino_Acid=="Deconjugated")%>%select(BA)%>%unique())$BA


#order and add bMCA as that's in the amino acid plots
decon_order<-c("CA","CDCA","DCA","LCA","aMCA","bMCA","HCA","HDCA","UDCA","7-keto DCA","EDCA","3-oxo LCA", "isoLCA","12-oxo LCA", "wMCA")

decon_stats<-MCBA_boxplot_stats%>%
            filter(qval<0.05 & Amino_Acid=="Deconjugated" & BA %in% decon_order)%>%
              mutate(BA=factor(BA, levels=decon_order))%>%
  mutate(x=as.numeric(BA))%>%
  mutate(xmin=x-0.2)%>%
  mutate(xmax=x+0.2)


ggplot(data=MCBA_boxplot%>%
         filter(Amino_Acid=="Deconjugated" & BA %in% decon_order)%>%
         mutate(BA=factor(BA, levels=decon_order)),
       aes(x=BA,y=value,fill=prepost, group=interaction(BA,prepost)))+
  geom_jitter(shape=21,
              position=position_jitterdodge(dodge.width=0.75,jitter.width=0.2),
              size=0.8,stroke=0.1)+
  geom_boxplot(alpha=0,linewidth=0.2)+
  theme_bw()+
  theme(text=element_text(size=10),
        legend.position="none",
        axis.title.y=element_markdown(),
        axis.text.x=element_text(angle=-90),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linewidth=0.2),
        panel.grid.minor= element_blank(),
        strip.background=element_blank())+
  labs(y = "",
       x= "",
       fill="Time Post FMT")+
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     breaks=c(0,1e2,1e4,1e6,1e8,1e10),
                     expand=c(0.03,0,0.1,0))+
  scale_fill_manual(values=anno_color$TIME.POINT.2)+
  stat_pvalue_manual(decon_stats,
                     label = "q.adj.signif",
                     size=2,
                     tip.length=0.01,
                     step.increase=0.07,
                     step.group.by="sort",
                     inherit.aes=F)

ggsave("Figures/MCBA/MCBA_Boxplots_Unconjugated.pdf",width=4.63, height=1.55)

###Now everything else for supp
ggplot(data=MCBA_boxplot%>%filter(!Amino_Acid %in% c(names(aa_names),"Deconjugated")), aes(x=Core,y=value,fill=prepost, group=interaction(Core,prepost)))+
  geom_jitter(shape=21,
              position=position_jitterdodge(dodge.width=0.75,jitter.width=0.2),
              size=2,stroke=0.25)+
  geom_boxplot(alpha=0,linewidth=0.2)+
  facet_wrap(~Amino_Acid,ncol=3)+
  theme_bw()+
  theme(text=element_text(size=10),
        legend.position="none",
        axis.title.y=element_markdown(),
        axis.text.x=element_text(angle=-90),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linewidth=0.2),
        panel.grid.minor= element_blank(),
        strip.background=element_blank())+
  labs(y = "MCBA Abundance",
       x= "",
       fill="Time Post FMT")+
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     limits=c(0,1e8),
                     breaks=c(0,1e2,1e4,1e6,1e8,1e10),
                     expand=c(0.03,0,0.1,0))+
  scale_fill_manual(values=anno_color$TIME.POINT.2)

ggsave("Figures/Supplementals/MCBA_Boxplots_notsig.pdf",width=8.25, height=5)

###Now decon for supp
ggplot(data=MCBA_boxplot%>%filter(Amino_Acid=="Deconjugated" & !BA %in% decon_order), aes(x=BA,y=value,fill=prepost, group=interaction(BA,prepost)))+
  geom_jitter(shape=21,
              position=position_jitterdodge(dodge.width=0.75,jitter.width=0.2),
              size=2,stroke=0.25)+
  geom_boxplot(alpha=0,linewidth=0.2)+
  theme_bw()+
  theme(text=element_text(size=10),
        legend.position="none",
        axis.title.y=element_markdown(),
        axis.text.x=element_text(angle=-90),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linewidth=0.2),
        panel.grid.minor= element_blank(),
        strip.background=element_blank())+
  labs(y = "BA Abundance",
       x= "",
       fill="Time Post FMT")+
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     limits=c(0,1e8),
                     breaks=c(0,1e2,1e4,1e6,1e8,1e10),
                     expand=c(0.03,0,0.1,0))+
  scale_fill_manual(values=anno_color$TIME.POINT.2)

ggsave("Figures/Supplementals/MCBA_Decon_Boxplots_notsig.pdf",width=3.2, height=2.5)
#width =2 for same axis spacing

```

# low vs high this is just a QC step
```{r}
baker_labels<-read.csv("Data and Metadata/MCBA_sampleID.csv")

MCBA_raw_low<-read.xlsx("Data and Metadata/20231210_Neg_BA_Updated_Full_IsomerIDed.xlsx",sheetIndex = 2)%>%
      dplyr::rename("BA"="Molecule.Name")%>%
      rename_with(~str_remove(.,".Total.Area.MS1"))%>%
      rename_at(baker_labels$Rank, ~ baker_labels$CLIENT.IDENTIFIER)%>%
      mutate(source="low")%>%
      gather(key="CLIENT.IDENTIFIER",value="value",-BA,-source)%>%
      mutate(BA=str_remove(BA,"\\*"))

lvh<-MCBA_raw_low%>%
    rbind(.,MCBA_raw%>%
                gather(key="CLIENT.IDENTIFIER",value="value",-BA)%>%
                filter(BA %in% MCBA_raw_low$BA)%>%
                mutate(source="high"))%>%
  group_by(source,BA)%>%
  arrange(-value)%>%
  mutate(rankorder=1:n())%>%
  mutate(permax=value/max(value))%>%
  mutate(value=if_else(value==0,NA,value))%>%
  mutate(medscale=value/median(value,na.rm=T))%>%
  mutate(value=if_else(is.na(value),0,value))%>%
  mutate(medscale=log10(if_else(is.na(medscale),0,medscale)))

v<-ggplot(lvh%>%select(-rankorder,-permax,-medscale)%>%spread(key="source",value="value"),aes(y=high,x=low,color=BA))+
  geom_point()+
  theme_bw()+
  scale_x_log10()+
  scale_y_log10()+
  ggtitle("Values")

m<-ggplot(lvh%>%select(-value,-rankorder,-permax)%>%spread(key="source",value="medscale"),aes(y=high,x=low,color=BA))+
  geom_point()+
  theme_bw()+
  ggtitle("scaled")

r<-ggplot(lvh%>%select(-value,-permax,-medscale)%>%spread(key="source",value="rankorder"),aes(y=high,x=low,color=BA))+
  geom_point()+
  theme_bw()+
  ggtitle("Rank order")

pm<-ggplot(lvh%>%select(-value,-rankorder,-medscale)%>%spread(key="source",value="permax"),aes(y=high,x=low,color=BA))+
  geom_point()+
  theme_bw()+
  ggtitle("percent max")

ggarrange(v,m,r,pm,nrow=2,ncol=2,common.legend = T)
ggsave("Figures/MCBA/low vs high check.pdf",width=7, height=6)

```

# Correlation to Pathway
```{r}
#Correlation to Pathway
pathway_for_cor<-pathway_raw%>%
  #filter(pathway %in% pathway_siglist$pathway)%>%
  filter(str_detect(pathway,"Amino-acid_biosynthesis"))%>%
  separate(pathway,into=c("type","amino","note"),sep=";",extra="merge")%>%
  separate(amino,into=c("amino","bolism"),sep="_",extra="merge")%>%
  filter(!str_detect(bolism,"regulation"))%>%
  select(amino,samplekey$CLIENT.IDENTIFIER)%>%
  gather(key="CLIENT.IDENTIFIER",value="CPM",-amino)%>%
  group_by(amino,CLIENT.IDENTIFIER)%>%
  summarize(CPM=sum(CPM))%>%
  spread(key="CLIENT.IDENTIFIER",value="CPM")%>%
  select(amino, samplekey$CLIENT.IDENTIFIER)%>%
  column_to_rownames("amino")%>%
  t()

pathway_for_cor<-pathway_raw%>%
  filter(str_detect(pathway,"Amino-acid_biosynthesis"))%>%
  filter(!str_detect(pathway,"regulation"))%>%
  column_to_rownames("pathway")%>%
  t()

source("Scripts/rm_corr_all.R")
rmcor_MCBA_path<-rmcorr_test(samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
                           "SUBJ.ID",
                           pathway_for_cor,
                           MCBA_for_cor,
                          "BH")

MCBA_path_cor_plot<-rmcor_MCBA_path%>%
  dplyr::rename("BA"=df2,"pathway"=df1)%>%
  #mutate(BA=factor(BA, levels=rev(MCBA_forest_plot$BA)))%>%
  filter(p.adj<=0.05)%>%
  select(-p,-p.adj)%>%
  #filter(!is.na(BA))%>%
  #complete(BA,fill=list(clade_name="drop",r=0))%>%
  spread(key=pathway,value=r,fill=0)%>%
  left_join(.,MCBA_metadata%>%select(BA,Amino_Acid),by="BA")%>%
  arrange(Amino_Acid)%>%
  select(-Amino_Acid)%>%
  column_to_rownames("BA")

pheatmap(MCBA_path_cor_plot,
         cluster_rows=F,
         cluster_cols=T,
         color=c(rev(colorRampPalette(c("grey88","white","dodgerblue3"),bias=1)(30)),colorRampPalette(c("grey88","white", "red3"),bias=1)(30)),
         border_color="grey20",
         breaks=seq(-1,1,by=2/60),
         fontsize = 14,
         fontsize_row = 10,
         drop_levels = T,
         filename="Figures/MCBA/MCBA_vs_aa_pathway_rmcor.pdf",
         width=8.5,
         height=11)
```

# Correlate to Amino Acids
I don't think this is helpful either
```{r}
###Correlation to AA
AA_for_cor<-bch_for_cor%>%
  select(c("lysine", "serine", "valine", "alanine",
           "histidine","glycine","leucine", "proline" ,
           "arginine","cysteine","tyrosine"  ,
           "aspartate","glutamate", "glutamine", "threonine",
           "asparagine","isoleucine", "methionine","tryptophan","phenylalanine",
           "taurine","ornithine","citrulline","hydroxyproline", #non proteinogenic amino acids
           "isovalerate (C5)", #oxidative leucine
           "alpha-hydroxyisocaproate",# reductive leucine
           "isocaproate (i6:0)", #reductive leucine
           "5-aminovalerate", # reductive proline
           "4-hydroxyphenylacetate", #oxidative tyrosine
           "p-cresol", # oxidative tyrosine
           "3-(4-hydroxyphenyl)propionate", # reductive tyrosine
           "indoleacetate", # oxidative Tryptophan
           "phenylacetate", # oxidative phenylalanine
           "4-imidazoleacetate", #oxidative histidine
           "3-phenylpropionate (hydrocinnamate)"))

source("Scripts/rm_corr_all.R")
rmcor_MCBA_path<-rmcorr_test(samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
                           "SUBJ.ID",
                           pathway_for_cor,
                           AA_for_cor,
                          "BH")

AA_path_cor_plot<-rmcor_MCBA_path%>%
  dplyr::rename("AA"=df2,"pathway"=df1)%>%
  #mutate(BA=factor(BA, levels=rev(MCBA_forest_plot$BA)))%>%
  filter(p.adj<=0.05)%>%
  select(-p,-p.adj)%>%
  #filter(!is.na(BA))%>%
  #complete(BA,fill=list(clade_name="drop",r=0))%>%
  spread(key=pathway,value=r,fill=0)%>%
  column_to_rownames("AA")

pheatmap(AA_path_cor_plot,
         cluster_rows=T,
         cluster_cols=T,
         color=c(rev(colorRampPalette(c("grey88","white","dodgerblue3"),bias=1)(30)),colorRampPalette(c("grey88","white", "red3"),bias=1)(30)),
         breaks=seq(from=-1,to=1,by=2/60),
         border_color="grey20",
         fontsize = 14,
         drop_levels = T,
         filename="Figures/MCBA/path_vs_AA_rmcor.pdf",
         width=14,
         height=17)

```

#Data export for Mendeley
```{r data export}
main_boxplot<-MCBA_boxplot%>%
         filter(Amino_Acid %in% names(aa_names)|BA %in% decon_order)
sup_boxplot<-MCBA_boxplot%>%
         filter(!Amino_Acid %in% names(aa_names)&!BA %in% decon_order)


write_xlsx(list("5A"=MCBA_forest_plot,
                "5A,D linear mixed model"=MCBA_siglist,
                "5B"=MCBA_by_time,
                "5C"=MCBA_fam_cor_plot,
                "5C linear model"=rmcor_MCBA_fam,
                "5D"=main_boxplot),
           "Excel Exports/Figure 5.xlsx")

write_xlsx(list("S5"=sup_boxplot),
           "Excel Exports/Figure S5.xlsx")

```

