---
title: "Metadata Comparisons"
author: "Sam McMillan"
date: "2024-01-24"
output: html_document
---

```{r setup, require=T, include=T, warning=F}
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(rstatix)
library(writexl)
library(Maaslin2)

sessionInfo()
```

```{r dataframes}
samplekey_rich<-readRDS("Dataframes/samplekey_rich.rds")
bch_raw<-readRDS("Dataframes/bch_raw.rds")
tax_raw<-readRDS("Dataframes/tax_raw.rds")
genes_raw<-readRDS("Dataframes/genes_raw.rds")
MCBA_raw<-readRDS("Dataframes/MCBA_raw.rds")
pathway_raw<-readRDS("Dataframes/pathway_raw.rds")

source("Scripts/Color codes.R")
```

# NMDS Dataframes
I first want to look at the differences between timepoints of all of the different datasets that I have. We'll start with NMDS. 
```{r NMDS}
#Distance matrices
tax_dist<-tax_raw%>%
  filter(str_detect(clade_name,"s__"))%>% #species only
  column_to_rownames("clade_name")%>%
  t()%>%
  vegdist(method="bray")

bch_dist<-bch_raw%>%
  column_to_rownames("BIOCHEMICAL")%>%
  t()%>%
  vegdist(method="bray")

genes_dist<-genes_raw%>%
  filter(!str_detect(genefamily,"\\|") & genefamily!="UNMAPPED")%>% #remove species information and unmapped genes
  column_to_rownames("genefamily")%>%
  t()%>%
  vegdist(method="bray")

MCBA_dist<-MCBA_raw%>%
  column_to_rownames("BA")%>%
  t()%>%
  vegdist(method="bray")
```

# Rich NMDS plots
We're concerned there might be differences in our data due to sex/antibiotic usage/etc. and we have the metadata from the surveys. NMDS serves as a nice way to broadly look at this.
```{r NMDS plots}
for (i in c("TIME.POINT.2", colnames(samplekey_rich[6:20]))){
a<-ggplot(data=bch_nmds%>%
         left_join(.,samplekey_rich,by=c("CLIENT.IDENTIFIER","SUBJ.ID","TIME.POINT.2","prepost")),
       aes_string(x="NMDS1",y="NMDS2",fill=i))+
  geom_point(shape=21,size=2,show.legend=T)+
    coord_equal()+
  theme_bw(base_size = fsize)+
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5))+
  labs(color="")+
    scale_y_continuous(limits=c(-1.5,1.5),breaks=c(-1.5,-1,-0.5,0,0.5,1,1.5))+
    scale_x_continuous(limits=c(-2,2),breaks=c(-2,-1,0,1,2))+
  ggtitle("Metabolites")

b<-ggplot(data=genes_nmds%>%
         left_join(.,samplekey_rich,by=c("CLIENT.IDENTIFIER","SUBJ.ID","TIME.POINT.2","prepost")),
       aes_string(x="NMDS1",y="NMDS2",fill=i))+
  geom_point(shape=21,size=2,show.legend=T)+
    coord_equal()+
  theme_bw(base_size = fsize)+
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5))+
  labs(color="")+
    scale_y_continuous(limits=c(-1.5,1.5),breaks=c(-1.5,-1,-0.5,0,0.5,1,1.5))+
    scale_x_continuous(limits=c(-2,2),breaks=c(-2,-1,0,1,2))+
  ggtitle("Genes")

c<-ggplot(data=tax_nmds%>%
         left_join(.,samplekey_rich,by=c("CLIENT.IDENTIFIER","SUBJ.ID","TIME.POINT.2","prepost")),
       aes_string(x="NMDS1",y="NMDS2",fill=i))+
  geom_point(shape=21,size=2,show.legend=T)+
    coord_equal()+
  theme_bw(base_size = fsize)+
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5))+
  labs(color="")+
    scale_y_continuous(limits=c(-1.5,1.5),breaks=c(-1.5,-1,-0.5,0,0.5,1,1.5))+
    scale_x_continuous(limits=c(-2,2),breaks=c(-2,-1,0,1,2))+
  ggtitle("Taxonomy")

d<-ggplot(data=MCBA_nmds%>%
         left_join(.,samplekey_rich,by=c("CLIENT.IDENTIFIER","SUBJ.ID","TIME.POINT.2","prepost")),
       aes_string(x="NMDS1",y="NMDS2",fill=i))+
  geom_point(shape=21,size=2,show.legend=T)+
    coord_equal()+
  theme_bw(base_size = fsize)+
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5))+
  labs(color="")+
    scale_y_continuous(limits=c(-1.5,1.5),breaks=c(-1.5,-1,-0.5,0,0.5,1,1.5))+
    scale_x_continuous(limits=c(-2,2),breaks=c(-2,-1,0,1,2))+
  ggtitle("MCBA")

l<-if(i=="FMT_response_ordinal"){"bottom"}else{"right"}
  
ggarrange(c,b,a,d, ncol=2,nrow=2, common.legend=T,legend=l)%>%
  annotate_figure(.,top=text_grob(i))

ggsave(paste0("Figures/NMDS_metadata/",i,".pdf"),height=4,width=5)

  }
```
Nothing pops out as impacting the data, especially when compared to the sampling timepoints. We can do some stats to make sure but we're definitely not appropriately powered for this.

#Adonis comparison with interactions
```{r}
adonis_sex<-function(dist){
set.seed(102309)
  tpadonis<-dist%>%
  as.matrix()%>%
  as.data.frame()%>%
  rownames_to_column("CLIENT.IDENTIFIER")%>%
  left_join(.,samplekey_rich,by="CLIENT.IDENTIFIER")%>%
  column_to_rownames("CLIENT.IDENTIFIER")%>%
  pairwise.adonis2(as.dist(.[1:45]) ~ TIME.POINT.2*sex, data = ., nperm = 1000, strata="SUBJ.ID")

  set.seed(102309)
  ppadonis<-dist%>%
  as.matrix()%>%
  as.data.frame()%>%
  rownames_to_column("CLIENT.IDENTIFIER")%>%
  left_join(.,samplekey_rich,by="CLIENT.IDENTIFIER")%>%
  column_to_rownames("CLIENT.IDENTIFIER")%>%
  adonis2(as.dist(.[1:45]) ~ sex*prepost, data = ., permutations = 1000, strata=.$SUBJ.ID)
 
 
  return(do.call("rbind",list(
          tpadonis[["Pre_vs_2 week"]]%>%as.data.frame()%>%rownames_to_column(".")%>%mutate(comp="Pre_vs_2 week"),
          tpadonis[["Pre_vs_2 month"]]%>%as.data.frame()%>%rownames_to_column(".")%>%mutate(comp="Pre_vs_2 month"),
          tpadonis[["Pre_vs_6 month"]]%>%as.data.frame()%>%rownames_to_column(".")%>%mutate(comp="Pre_vs_6 month"),
          ppadonis%>%as.data.frame()%>%rownames_to_column(".")%>%mutate(comp="Pre vs Post")
          )))
          
}

bch<-adonis_sex(bch_dist)
tax<-adonis_sex(tax_dist)
MCBA<-adonis_sex(MCBA_dist)
genes<-adonis_sex(genes_dist)

write_xlsx(list("Untargeted Metabolomics"=bch,
                "Targeted Bile Acids"=MCBA,
                "Species"=tax,
                "Microbial Genes"=genes),
           "Excel Exports/Sex comparison_interactions.xlsx")
```


```{r maaslin check}
tax_fam<-tax_raw%>%
        filter(str_detect(clade_name,"\\;f\\_\\_")&!str_detect(clade_name,"\\;g\\_\\_"))%>%
        mutate(clade_name=str_extract(clade_name, "\\;f\\_\\_.*"))%>%
        mutate(clade_name=str_remove(clade_name,"\\;f\\_\\_"))

tax_spec<-tax_raw%>%
        filter(str_detect(clade_name,"\\;s\\_\\_"))

for (d in c("bch_raw","MCBA_raw","tax_fam","tax_spec","pathway_raw")){
data<-get(d)%>%
               dplyr::rename("feature"=1)%>%
               filter(!str_detect(feature, "\\|"))%>%
               mutate(feature=make_clean_names(feature))%>%
               column_to_rownames("feature")%>%
               mutate_all(~as.numeric(.))%>%
               t()

capture.output(Maaslin2(input_data = data,
             input_metadata = samplekey_rich%>%column_to_rownames("CLIENT.IDENTIFIER"),
             output = paste0("Maaslin output/",d,"_sex"),
             fixed_effects = c("sex"), 
             random_effects = "SUBJ.ID",
             reference=c("sex,Female"),
             plot_scatter=F,
             plot_heatmap = F,
             normalization="NONE",
             transform="LOG",
             analysis_method ="LM",
             correction="bonferroni",
             max_significance = 0.05), file=paste0("Maaslin output/",d,"_sex/console_output.txt")) 
}


all<-matrix(nrow=0,ncol=9,dimnames = list(NULL,c("feature","metadata","value","coef","stderr","N","N.not.0","pval","qval")))

for (d in c("bch_raw","MCBA_raw","tax_fam","tax_spec","pathway_raw")){
  data<-left_join(read_tsv(paste0("Maaslin output/",d,"_sex/significant_results.tsv")),get(d)%>%
                             select(1)%>%
                             dplyr::rename("name"=1)%>%
                             mutate(feature=make_clean_names(name))%>%
                             mutate(data=d),by="feature")
all<-rbind(all,data)
}

write.csv(all,"Excel Exports/sex_analysis.csv")

```

