---
title: "Reviewer Comments"
author: "Sam McMillan"
date: "2024-06-04"
output: html_document
---

```{r setup, require=T, include=T}
library(tidyverse)
library(pheatmap)
library(randomForest)
library(ggtext)
library(janitor)
library(writexl)
library(Maaslin2)

sessionInfo()
```

#Get Dataframes
```{r}
samplekey<-readRDS("Dataframes/samplekey.rds")%>%
  left_join(.,read.csv("Data and Metadata/abx pre metadata.csv"),by="SUBJ.ID")
fam_for_cor<-readRDS("Dataframes/fam_for_cor.rds")
bch_for_cor<-readRDS("Dataframes/bch_for_cor.rds")
MCBA_for_cor<-readRDS("Dataframes/MCBA_for_cor.rds")
top_bch_forest<-readRDS("Dataframes/top_bch_forest.rds")


bch_raw<-readRDS("Dataframes/bch_raw.rds")
MCBA_raw<-readRDS("Dataframes/MCBA_raw.rds")
tax_raw<-readRDS("Dataframes/tax_raw.rds")
MCBA_forest<-readRDS("Dataframes/MCBA order.rds")
bch_siglist<-readRDS("Dataframes/Stats/bch_siglist.rds")
MCBA_siglist<-readRDS("Dataframes/Stats/MCBA_siglist.rds")
pathway_siglist<-readRDS("Dataframes/Stats/pathway_siglist.rds")
fam_siglist<-readRDS("Dataframes/Stats/fam_siglist.rds")
spec_siglist<-readRDS("Dataframes/Stats/spec_siglist.rds")


source("Scripts/Color codes.R")
```

```{r}
source("Scripts/rm_corr_all.R")

rmcor_BCH_FAM<-rmcorr_test(samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
                           "SUBJ.ID",
                           bch_for_cor[top_bch_forest],
                          fam_for_cor[levels(fam_siglist$clade_name)],
                          "BH")

bch_vs_fam_top<-rmcor_BCH_FAM%>%
              arrange(r)%>%
              mutate(rank=1:n())%>%
              filter(rank<=10|rank>max(rank-10))

bchvfam_plotdata<-bch_for_cor[top_bch_forest]%>%
  rownames_to_column("CLIENT.IDENTIFIER")%>%
  gather(key="BIOCHEMICAL",value="value",-CLIENT.IDENTIFIER)%>%
  filter(str_detect(BIOCHEMICAL,"carnitine"))%>%
  left_join(.,samplekey,by="CLIENT.IDENTIFIER")%>%
  left_join(.,fam_for_cor["Enterobacteriaceae"]%>%rownames_to_column("CLIENT.IDENTIFIER"),by="CLIENT.IDENTIFIER")

ggplot(data=bchvfam_plotdata,aes(x=log10(value),y=Enterobacteriaceae,fill=TIME.POINT.2,group=SUBJ.ID))+
  geom_line(color="black",alpha=0.1)+
  geom_point(shape=21, size=2)+
  facet_wrap(~BIOCHEMICAL,scales="free",ncol=2)+
  theme_bw()+
  theme(strip.background=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor= element_blank())+
  labs(x="log10(Metabolite Abundance)",fill="", y="Relative Abundance Enterobacteriaceae")+
  scale_fill_manual(values=anno_color$TIME.POINT.2)

ggsave("Revision Outputs/Entero_vs_acylcarn_scatter.pdf", height=4, width=6, units="in")

```
# Reviewers were concerned about the impact of antibiotic treatment pre and post FMT
I put together all of the data we had on this. A lot is unknown

# Linear Modeling 
Data is already normalized: Median set to 1 for metabolites,Relative abundance for taxonomy, Counts per million for genes
Linear modelling is great because we have paired longitudinal samples (SUBJ.ID), but we also want to look at pre vs post (multiple post per pre). 
```{r linmod, message=F,warning=F}
tax_fam<-tax_raw%>%
        filter(str_detect(clade_name,"\\;f\\_\\_")&!str_detect(clade_name,"\\;g\\_\\_"))%>%
        mutate(clade_name=str_extract(clade_name, "\\;f\\_\\_.*"))%>%
        mutate(clade_name=str_remove(clade_name,"\\;f\\_\\_"))

tax_spec<-tax_raw%>%
        filter(str_detect(clade_name,"\\;s\\_\\_"))

#Not species level
for (d in c("bch_raw","MCBA_raw","tax_fam","tax_spec")){
  
  data<-get(d)%>%
               dplyr::rename("feature"=1)%>%
               filter(!str_detect(feature, "\\|"))%>%
               mutate(feature=make_clean_names(feature))%>%
               column_to_rownames("feature")%>%
               mutate_all(~as.numeric(.))%>%
               t()
  
    dir.create(paste0("Maaslin output/",d,"_prepost_abx"))
    dir.create(paste0("Maaslin output/",d,"_TIME.POINT.2_abx"))
  
    capture.output(Maaslin2(input_data = data,
             input_metadata = samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
             output = paste0("Maaslin output/",d,"_prepost_abx"),
             fixed_effects = c("prepost","pre_Abx"), 
             random_effects = "SUBJ.ID",
             reference=c("prepost,Pre","pre_Abx,Yes"),
             plot_scatter=F,
             plot_heatmap = F,
             normalization="NONE",
             transform="LOG",
             analysis_method ="LM",
             correction="bonferroni",
             max_significance = 0.05), file=paste0("Maaslin output/",d,"_prepost_abx/console_output.txt")) 
    
    capture.output(Maaslin2(input_data = data,
             input_metadata = samplekey%>%column_to_rownames("CLIENT.IDENTIFIER"),
             output = paste0("Maaslin output/",d,"_TIME.POINT.2_abx"),
             fixed_effects = c("TIME.POINT.2","pre_Abx"), 
             random_effects = "SUBJ.ID",
             reference=c("TIME.POINT.2,Pre","pre_Abx,Yes"),
             plot_scatter=F,
             plot_heatmap = F,
             normalization="NONE",
             transform="LOG",
             analysis_method ="LM",
             correction="bonferroni",
             max_significance = 0.05), file=paste0("Maaslin output/",d,"_TIME.POINT.2_abx/console_output.txt"))  
}
```

# Organize lists of significant metabolites/families
Put the pre vs post comparisons with the pre vs each time point comparisons and save them for later reference
```{r siglists}
bch_siglist_abx<-rbind(read_tsv("Maaslin output/bch_raw_prepost_abx/all_results.tsv"),
                   read_tsv("Maaslin output/bch_raw_TIME.POINT.2_abx/all_results.tsv"))%>%
            left_join(.,bch_raw%>%select(BIOCHEMICAL)%>%mutate(feature=make_clean_names(BIOCHEMICAL)), by="feature")%>%
            filter(qval<=0.05)%>%
            mutate(dir=case_when(coef>0~"Up",coef<0~"down"))%>%
            mutate(BIOCHEMICAL=factor(BIOCHEMICAL,levels=unique(.$BIOCHEMICAL)))

MCBA_siglist_abx<-rbind(read_tsv("Maaslin output/MCBA_raw_prepost_abx/all_results.tsv"),
                   read_tsv("Maaslin output/MCBA_raw_TIME.POINT.2_abx/all_results.tsv"))%>%
            left_join(.,MCBA_raw%>%select(BA)%>%mutate(feature=make_clean_names(BA)), by="feature")%>%
            filter(qval<=0.05)%>%
            mutate(dir=case_when(coef>0~"Up",coef<0~"down"))%>%
            mutate(BA=factor(BA,levels=unique(.$BA)))

fam_siglist_abx<-rbind(read_tsv("Maaslin output/tax_fam_prepost_abx/all_results.tsv"),
                   read_tsv("Maaslin output/tax_fam_TIME.POINT.2_abx/all_results.tsv"))%>%
            left_join(.,tax_fam%>%select(clade_name)%>%mutate(feature=make_clean_names(clade_name)), by="feature")%>%
            filter(qval<=0.05)%>%
            mutate(dir=case_when(coef>0~"Up",coef<0~"down"))%>%
            mutate(clade_name=factor(clade_name,levels=unique(.$clade_name)))

spec_siglist_abx<-rbind(read_tsv("Maaslin output/tax_spec_prepost_abx/all_results.tsv"),
                   read_tsv("Maaslin output/tax_spec_TIME.POINT.2_abx/all_results.tsv"))%>%
            left_join(.,tax_spec%>%select(clade_name)%>%mutate(feature=make_clean_names(clade_name)), by="feature")%>%
            filter(qval<=0.05)%>%
            mutate(dir=case_when(coef>0~"Up",coef<0~"down"))%>%
            mutate(clade_name=factor(clade_name,levels=unique(.$clade_name)))

```
# Check for changes
```{r}

bch_check<-filter(bch_siglist_abx,BIOCHEMICAL %in% top_bch_forest)

bch_missing<-top_bch_forest[-which(top_bch_forest %in% filter(bch_siglist_abx,BIOCHEMICAL %in% top_bch_forest)$BIOCHEMICAL)]

```

